Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
changed soi=> inf

\newcommand{\phirs}{\frac{G \Menc}{\rs}}
\newcommand{\soi}{\rm soi}
\newcommand{\rsoi}{r_{\soi}}
\newcommand{\rIa}{r_{\rm Ia}}
\newcommand{\EIa}{E_{\rm Ia}}
\newcommand{\RateIa}{R_{\rm Ia}}
\newcommand{\sigsoi}{\sigma_{\soi}}

@@

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & black hole mass \\
$\tilde{v}_{w}$ & total heating parameter, including minimum heating
rate from stellar and black hole velocity dispersion \\
$v_{w}$ & total heating parameter, excluding heating rate from
velocity dispersion \\
$\sigma_0$ & stellar velocity dispersion--assumed to be radially constant.\\
$\zeta$ & heating parameter, $\zeta \equiv \sqrt{1 + (v_w/\sigma_0)^2}$ (eq.~[\ref{eq:zetacrit}])\\
$r_{s}$ & stagnation radius (where gas radial velocity goes to zero) \\
$r_{\rm soi}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\ 
$r_{\rm Ia}$ & radius interior to which Ia SN are infrequent compared to the dynamical/inflow timescale (eq.~[\ref{eq:rIa}]) \\ 
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar surface brightness
profile interior to the break radius \\
$\delta$ & power-law slope of the 3d stellar density profile inside of
the break radius. $\delta=\Gamma+1$.\\
$\dot{M}$ & mass accretion rate  \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & `equilibrium' accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$t_{\rm h}$ & Hubble timescale \\
$\nu$ & gas density power-law slope at the stagnation radius
(eq.~[\ref{eq:densSlope}]) \\
\hline
\label{table:definitions}  
\end{tabular}
%\begin{tablenotes}
%\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\label{sec:gal_model}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to
measure the radial surface brightness profiles for hundreds of nearby
early type galaxies. The measured brightness profile is generally
well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law
slope, $\Gamma$, to an outer power law slope, $\beta > \Gamma$, at a
break radius, $\rb$.  Assuming the stellar population is spherically
symmetric, this corresponds to a stellar density $\rhostar \propto
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for
$r \gg \rb$.  To good approximation
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star|_{\rsoi} \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star|_{r_b} \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star|_{\rsoi}$ is the stellar density at the radius of the black hole sphere of influence (soi), 
\begin{equation}
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\end{equation}
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the
second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$
relationship of \citet{McConnellMa+:2011a},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
Though of questionable validity for low mass black holes (e.g., \citealt{Greene&Ho07}), we adopt equation (\ref{eq:Msigma}) for lack of a better alternative.

@@

Equation (\ref{eq:mdot_gas}) closely resembles the Bondi formula (eq.~[\ref{eq:bondi}]) provided that $r_{\rm B}$ is replaced by $\rs$.  Indeed, for $r_{s} < r_{\rm soi}$ we have that (eq.~[\ref{eq:stag_simple}])
\begin{align}
  \rs\approx\frac{13+8\Gamma}{4+2\Gamma}\frac{G \Mbh}{\densSlope
    \vw^2} \approx \frac{13+8\Gamma}{12+6\Gamma}r_{\rm B},
  \label{eq:rbondi}
\end{align}
where the second equality makes use of equation (\ref{eq:Tanalytic}).

@@

Our analytic derivations neglect the effects of heat conduction, an
assumption we now check.  The ratio of the magnitude of the conductive
heating rate (eq.~[\ref{eq:qdotcond}]) to the external heating rate at
the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}}
\tilde{v}_{w}^{2}}
 \left(1+\frac{\kappa_0 T^{7/2}|_{r_{\rm s}}}{5 \rs
    \phi \rho c_s^3}\right)^{-1}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 30 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  2 \phi & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}),
(\ref{eq:stag_simple}), and we have made the approximations $\nabla^{2} \sim
1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  The
stellar density profile is approximated as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    7.9 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.3 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker
break radius.  Equation (\ref{eq:conduction}) shows that, even when conduction is saturated, our neglect of heat conduction
results in at most an order unity correction for plausible, causal values of
the saturation parameter $\phi < 0.1$.  Our numerical experiments which include conductive heating confirm this
($\S\ref{sec:numerical}$).

@@

Based on our numerical results (\ref{sec:numerical}) and the work of \citet{McCourt+12}) we define thermally stable flows according to the criterion $\dot{q}_{\rm heat} > 10|\dot{q}_{\rm rad}|$ ($t_{\rm
  cool} > 10t_{\rm ff}$) being satisfied near the stagnation radius.  This condition translates into a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  {\rm max}\begin{cases}
   300 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   260 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}  \\
\approx \sigma_0(r_{b}/r_{\rm soi})^{0.5(1-\Gamma)} &, \zeta \ll \zeta_c, 
  \end{cases}
  \label{eq:cooling3}
\end{align}
where the final condition corresponds to the low wind heating limit
($\zeta \ll \zeta_c$; eq.~[\ref{eq:zetacrit}]), for which our
numerical and analytic results show that the stagnation radius moves
out to the break radius $r_b$; the quasi-hydrostatic structure that
develops in this case greatly increases the gas density, resulting in
thermal instability.

The minimum heating rate for thermal stability (eq.~[\ref{eq:cooling2}]) corresponds to maximum thermally-stable accretion rate,
\begin{align}
  \frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq {\rm
    min}\begin{cases}
    2\times 10^{-4} \eta_{0.02}^{0.47} M_{\bullet,8}^{0.36} &, \text{core}, \zeta > \zeta_c\\
    4\times 10^{-6}\eta_{0.02}M_{\bullet,8}^{-0.37} &, \text{core}, \zeta < \zeta_c \\
    7\times 10^{-5} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.28}    &, \text{cusp}, \zeta > \zeta_c\\
    1\times 10^{-4}\eta_{0.02}&, \text{cusp}, \zeta < \zeta_c
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
where the first line for each case is derived using equation
(\ref{eq:eddr_analytic}) to calculate $\dot{M}(v_{\rm TI})$, whereas
the second line corresponds to the low wind heating limit $\zeta \ll
\zeta_c$.  $\zeta_c$ is calculated using the empirical correlation for
core galaxies $r_{b}/r_{\rm soi} \approx 7.6M_{\bullet,8}^{-0.2}$ which
follows from equation (\ref{eq:rsoi}) and the $r_b-M_{\bullet}$
correlation derived from the results of \citet{LauerFaber+:2007a}.

@@

  \citet{EmsellemCappellari+:2007a} use two-dimensional kinematic data
  to measure the ratio of ordered to random motion in a sample of
  early type galaxies, which they quantify at each galactic radius $R$
  by the parameter
  \begin{align}
    \lambda_R \equiv \frac{\langle r|V|\rangle}{\langle R\sqrt{V^2+\sigma^2}\rangle} \underset{R \ll r_{\rm soi}}\sim \frac{V_{\phi}}{\sigma},
  \end{align}
  where $\sigma$ is the velocity dispersion and the brackets indicate
  a luminosity-weighted average.  The circularization radius of the
  accretion flow can be written in terms of $\lambda_R$ as

\begin{equation}
\frac{r_{\rm circ}}{r_{s}} \approx \frac{r_{s}}{r_{\rm soi}}\lambda_{R}^{2} \lesssim \lambda_{R}^{2},
\label{eq:rcirc}
\end{equation}
where we have used the definition $r_{\rm soi} \equiv
GM_{\bullet}/\sigma^{2}$ and in the second inequality have assumed
that $r_{\rm s} \lesssim r_{\rm soi}$, a condition which is satisfied
for the thermally-stable solutions of interest.

@@

  {$M_{\bullet}$} & {$v_{w}$} & {$r_b^{(b)}$} &   $r_{\rm s}/r_{\rm soi}$ &  {$\left(\frac{\dot{M}}{\dot{M}_{\rm edd}}\frac{\eta}{0.2}\right)_{r_{s}}^{(c)}$} &

@@

Figure~\ref{fig:stag} shows our calculation of the stagnation radius
$r_{\rm s}/r_{\rm soi}$ as a function of the wind heating parameter
$\zeta = \sqrt{1+(v_w/\sigma_0)^{2}}$, with different colors
showing different values of $v_{w}$.  Cusp and core galaxies are
marked with square and triangles, respectively.  Shown for comparison
are our analytic results (eq.~[\ref{eq:rs2main}]) with solid and
dashed lines for cusp and core galaxies, respectively, calculated
assuming $\densSlope = 1$ and $\densSlope= 0.6$, respectively
(eq.~[\ref{eq:densSlope}]).

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm soi}$).  However, for sufficiently low heating the
stagnation radius diverges above the analytic estimate, approaching
the stellar break radius $r_b \gg r_{\rm soi}$.  This divergence
occurs approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which in Appendix
\ref{app:be} is shown to coincide with the Bernoulli parameter of the gas decreasing below unity near the stagnation radius.  This divergence of the stagnation radius to large radii hence occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.8$) with large break radii.  This dependence explains the behavior of the two core galaxies, shown as
orange triangles vertically aligned at the same value of
$\zeta$; the solutions correspond to the same black hole mass
($\Mbh = 10^8 \Msun$) and heating rate ($v_w = 600$ km s$^{-1}$) but are calculated assuming
different values of the break radius: $r_{b} =$ 25, 50 pc (from top to
bottom).

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} Stagnation radius $r_{s}$ in units of the
    sphere of influence radius $r_{\rm soi}$ (eq.~[\ref{eq:rsoi}]) for
    galaxies in our sample as a function of the stellar wind heating
    parameter $\zeta \equiv \sqrt{1+(v_w/\sigma_0)^{2}}$.
    Green, orange, and blue symbols correspond to different values of
    $v_{w} =$ 300, 600, and 1200 km s$^{-1}$, respectively.  Squares
    correspond to cusp galaxies ($\Gamma = 0.8$), while triangles
    correspond to cores ($\Gamma = 0.1$). Green crosses correspond to
    cusp solutions which would be thermally unstable. The black curves
    correspond to the analytic prediction from equation
    (\ref{eq:rs2main}), with solid and dashed curves calculated for
    parameters $(\Gamma=0.8, \densSlope\simeq1)$ and
    $(\Gamma=0.1,\densSlope\simeq0.6)$, respectively. }
\end{figure}

@@

The left panel of Figure \ref{fig:miller} is calculated assuming a
constant low value for the radiative efficiency of $\epsilon_X =
10^{-4}$, typical of those estimated for low luminosity AGN (e.g.,
\citealt{Ho:2009a}), while the right panel luminosities are calculated
instead assuming an efficiency of $\epsilon_X= 10 \epsilon_{\rm
  bol}\dot{M}/\dot{M}_{\rm edd}$ that depends on the Eddington ratio
as predicted by some models for radiatively inefficient accretion
flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}).  Shown for
comparison are the X-ray measurements (green stars) and upper limits
(gray triangles) from the sample of early-type galaxies compiled by
\citet{Miller+15} (cf.~\citealt{Gallo+10}).  A green line shows the
best power-law fit to the X-ray luminosity from \citet{Miller+15},
given by $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ with
$\alpha = -0.2$ (see also \citep{Zhang+09, Pellegrini10, Gallo+10}).
Also shown are the maximum accretion rates, respectively, for
thermally stable accretion (eq.~[\ref{eq:Mdotmax}]; yellow), as set by
SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]; teal), and as allowed by
Compton heating feedback (eq.~[\ref{eq:MdotC}]; red).

>>>>

\newcommand{\phirs}{\frac{G \Menc}{\rs}}
\newcommand{\soi}{\rm inf}
\newcommand{\rsoi}{r_{\soi}}
\newcommand{\rIa}{r_{\rm Ia}}
\newcommand{\EIa}{E_{\rm Ia}}
\newcommand{\RateIa}{R_{\rm Ia}}
\newcommand{\sigsoi}{\sigma_{\soi}}

@@

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & black hole mass \\
$\tilde{v}_{w}$ & total heating parameter, including minimum heating
rate from stellar and black hole velocity dispersion \\
$v_{w}$ & total heating parameter, excluding heating rate from
velocity dispersion \\
$\sigma_0$ & stellar velocity dispersion--assumed to be radially constant.\\
$\zeta$ & heating parameter, $\zeta \equiv \sqrt{1 + (v_w/\sigma_0)^2}$ (eq.~[\ref{eq:zetacrit}])\\
$r_{s}$ & stagnation radius (where gas radial velocity goes to zero) \\
$r_{\rm inf}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\ 
$r_{\rm Ia}$ & radius interior to which Ia SN are infrequent compared to the dynamical/inflow timescale (eq.~[\ref{eq:rIa}]) \\ 
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar surface brightness
profile interior to the break radius \\
$\delta$ & power-law slope of the 3d stellar density profile inside of
the break radius. $\delta=\Gamma+1$.\\
$\dot{M}$ & mass accretion rate  \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & `equilibrium' accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$t_{\rm h}$ & Hubble timescale \\
$\nu$ & gas density power-law slope at the stagnation radius
(eq.~[\ref{eq:densSlope}]) \\
\hline
\label{table:definitions}  
\end{tabular}
%\begin{tablenotes}
%\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\label{sec:gal_model}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to
measure the radial surface brightness profiles for hundreds of nearby
early type galaxies. The measured brightness profile is generally
well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law
slope, $\Gamma$, to an outer power law slope, $\beta > \Gamma$, at a
break radius, $\rb$.  Assuming the stellar population is spherically
symmetric, this corresponds to a stellar density $\rhostar \propto
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for
$r \gg \rb$.  To good approximation
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star|_{\rsoi} \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star|_{r_b} \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star|_{\rsoi}$ is the stellar density at the radius of the black hole sphere of influence (inf), 
\begin{equation}
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\end{equation}
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the
second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$
relationship of \citet{McConnellMa+:2011a},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
Though of questionable validity for low mass black holes (e.g., \citealt{Greene&Ho07}), we adopt equation (\ref{eq:Msigma}) for lack of a better alternative.

@@

Equation (\ref{eq:mdot_gas}) closely resembles the Bondi formula (eq.~[\ref{eq:bondi}]) provided that $r_{\rm B}$ is replaced by $\rs$.  Indeed, for $r_{s} < r_{\rm inf}$ we have that (eq.~[\ref{eq:stag_simple}])
\begin{align}
  \rs\approx\frac{13+8\Gamma}{4+2\Gamma}\frac{G \Mbh}{\densSlope
    \vw^2} \approx \frac{13+8\Gamma}{12+6\Gamma}r_{\rm B},
  \label{eq:rbondi}
\end{align}
where the second equality makes use of equation (\ref{eq:Tanalytic}).

@@

Our analytic derivations neglect the effects of heat conduction, an
assumption we now check.  The ratio of the magnitude of the conductive
heating rate (eq.~[\ref{eq:qdotcond}]) to the external heating rate at
the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}}
\tilde{v}_{w}^{2}}
 \left(1+\frac{\kappa_0 T^{7/2}|_{r_{\rm s}}}{5 \rs
    \phi \rho c_s^3}\right)^{-1}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 30 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  2 \phi & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}),
(\ref{eq:stag_simple}), and we have made the approximations $\nabla^{2} \sim
1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  The
stellar density profile is approximated as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm inf}^{3}}\left(\frac{r_{\rm s}}{r_{\rm inf}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    7.9 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.3 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker
break radius.  Equation (\ref{eq:conduction}) shows that, even when conduction is saturated, our neglect of heat conduction
results in at most an order unity correction for plausible, causal values of
the saturation parameter $\phi < 0.1$.  Our numerical experiments which include conductive heating confirm this
($\S\ref{sec:numerical}$).

@@

Based on our numerical results (\ref{sec:numerical}) and the work of \citet{McCourt+12}) we define thermally stable flows according to the criterion $\dot{q}_{\rm heat} > 10|\dot{q}_{\rm rad}|$ ($t_{\rm
  cool} > 10t_{\rm ff}$) being satisfied near the stagnation radius.  This condition translates into a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  {\rm max}\begin{cases}
   300 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   260 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}  \\
\approx \sigma_0(r_{b}/r_{\rm inf})^{0.5(1-\Gamma)} &, \zeta \ll \zeta_c, 
  \end{cases}
  \label{eq:cooling3}
\end{align}
where the final condition corresponds to the low wind heating limit
($\zeta \ll \zeta_c$; eq.~[\ref{eq:zetacrit}]), for which our
numerical and analytic results show that the stagnation radius moves
out to the break radius $r_b$; the quasi-hydrostatic structure that
develops in this case greatly increases the gas density, resulting in
thermal instability.

The minimum heating rate for thermal stability (eq.~[\ref{eq:cooling2}]) corresponds to maximum thermally-stable accretion rate,
\begin{align}
  \frac{\dot{M}_{\rm TI}}{\dot{M}_{\rm edd}} \simeq {\rm
    min}\begin{cases}
    2\times 10^{-4} \eta_{0.02}^{0.47} M_{\bullet,8}^{0.36} &, \text{core}, \zeta > \zeta_c\\
    4\times 10^{-6}\eta_{0.02}M_{\bullet,8}^{-0.37} &, \text{core}, \zeta < \zeta_c \\
    7\times 10^{-5} \eta_{0.02}^{0.59} M_{\bullet,8}^{0.28}    &, \text{cusp}, \zeta > \zeta_c\\
    1\times 10^{-4}\eta_{0.02}&, \text{cusp}, \zeta < \zeta_c
  \end{cases}
  \label{eq:Mdotmax}
\end{align}
where the first line for each case is derived using equation
(\ref{eq:eddr_analytic}) to calculate $\dot{M}(v_{\rm TI})$, whereas
the second line corresponds to the low wind heating limit $\zeta \ll
\zeta_c$.  $\zeta_c$ is calculated using the empirical correlation for
core galaxies $r_{b}/r_{\rm inf} \approx 7.6M_{\bullet,8}^{-0.2}$ which
follows from equation (\ref{eq:rsoi}) and the $r_b-M_{\bullet}$
correlation derived from the results of \citet{LauerFaber+:2007a}.

@@

  \citet{EmsellemCappellari+:2007a} use two-dimensional kinematic data
  to measure the ratio of ordered to random motion in a sample of
  early type galaxies, which they quantify at each galactic radius $R$
  by the parameter
  \begin{align}
    \lambda_R \equiv \frac{\langle r|V|\rangle}{\langle R\sqrt{V^2+\sigma^2}\rangle} \underset{R \ll r_{\rm inf}}\sim \frac{V_{\phi}}{\sigma},
  \end{align}
  where $\sigma$ is the velocity dispersion and the brackets indicate
  a luminosity-weighted average.  The circularization radius of the
  accretion flow can be written in terms of $\lambda_R$ as

\begin{equation}
\frac{r_{\rm circ}}{r_{s}} \approx \frac{r_{s}}{r_{\rm inf}}\lambda_{R}^{2} \lesssim \lambda_{R}^{2},
\label{eq:rcirc}
\end{equation}
where we have used the definition $r_{\rm inf} \equiv
GM_{\bullet}/\sigma^{2}$ and in the second inequality have assumed
that $r_{\rm s} \lesssim r_{\rm inf}$, a condition which is satisfied
for the thermally-stable solutions of interest.

@@

  {$M_{\bullet}$} & {$v_{w}$} & {$r_b^{(b)}$} &   $r_{\rm s}/r_{\rm inf}$ &  {$\left(\frac{\dot{M}}{\dot{M}_{\rm edd}}\frac{\eta}{0.2}\right)_{r_{s}}^{(c)}$} &

@@

Figure~\ref{fig:stag} shows our calculation of the stagnation radius
$r_{\rm s}/r_{\rm inf}$ as a function of the wind heating parameter
$\zeta = \sqrt{1+(v_w/\sigma_0)^{2}}$, with different colors
showing different values of $v_{w}$.  Cusp and core galaxies are
marked with square and triangles, respectively.  Shown for comparison
are our analytic results (eq.~[\ref{eq:rs2main}]) with solid and
dashed lines for cusp and core galaxies, respectively, calculated
assuming $\densSlope = 1$ and $\densSlope= 0.6$, respectively
(eq.~[\ref{eq:densSlope}]).

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm inf}$).  However, for sufficiently low heating the
stagnation radius diverges above the analytic estimate, approaching
the stellar break radius $r_b \gg r_{\rm inf}$.  This divergence
occurs approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which in Appendix
\ref{app:be} is shown to coincide with the Bernoulli parameter of the gas decreasing below unity near the stagnation radius.  This divergence of the stagnation radius to large radii hence occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.8$) with large break radii.  This dependence explains the behavior of the two core galaxies, shown as
orange triangles vertically aligned at the same value of
$\zeta$; the solutions correspond to the same black hole mass
($\Mbh = 10^8 \Msun$) and heating rate ($v_w = 600$ km s$^{-1}$) but are calculated assuming
different values of the break radius: $r_{b} =$ 25, 50 pc (from top to
bottom).

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} Stagnation radius $r_{s}$ in units of the
    sphere of influence radius $r_{\rm inf}$ (eq.~[\ref{eq:rsoi}]) for
    galaxies in our sample as a function of the stellar wind heating
    parameter $\zeta \equiv \sqrt{1+(v_w/\sigma_0)^{2}}$.
    Green, orange, and blue symbols correspond to different values of
    $v_{w} =$ 300, 600, and 1200 km s$^{-1}$, respectively.  Squares
    correspond to cusp galaxies ($\Gamma = 0.8$), while triangles
    correspond to cores ($\Gamma = 0.1$). Green crosses correspond to
    cusp solutions which would be thermally unstable. The black curves
    correspond to the analytic prediction from equation
    (\ref{eq:rs2main}), with solid and dashed curves calculated for
    parameters $(\Gamma=0.8, \densSlope\simeq1)$ and
    $(\Gamma=0.1,\densSlope\simeq0.6)$, respectively. }
\end{figure}

@@

The left panel of Figure \ref{fig:miller} is calculated assuming a
constant low value for the radiative efficiency of $\epsilon_X =
10^{-4}$, typical of those estimated for low luminosity AGN (e.g.,
\citealt{Ho:2009a}), while the right panel luminosities are calculated
instead assuming an efficiency of $\epsilon_X= 10 \epsilon_{\rm
  bol}\dot{M}/\dot{M}_{\rm edd}$ that depends on the Eddington ratio
as predicted by some models for radiatively inefficient accretion
flows (\citealt{Narayan&Yi95}; \citealt{Narayan+98}).  Shown for
comparison are the X-ray measurements (green stars) and upper limits
(gray triangles) from the sample of early-type galaxies compiled by
\citet{Miller+15} (cf.~\citealt{Gallo+10}).  A green line shows the
best power-law fit to the X-ray luminosity from \citet{Miller+15},
given by $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ with
$\alpha = -0.2$ (see also \citealt{Zhang+09, Pellegrini10, Gallo+10}).
Also shown are the maximum accretion rates, respectively, for
thermally stable accretion (eq.~[\ref{eq:Mdotmax}]; yellow), as set by
SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]; teal), and as allowed by
Compton heating feedback (eq.~[\ref{eq:MdotC}]; red).
